simulate Hospitalizations

### 2023-01-26 5_roadmaps_future_projects/RSV_maternal_vac_eff/Antibody_Sim.Rmd
add materials

---
title: "age structure"
author: "ZHE ZHENG, Dan Weinberger, modified from Matlab code by Virginia Pitzer"
date: "12/26/2020"
output: html_document
---
```{r}
library(deSolve)
library(RColorBrewer)
library(reshape2)
```
In Pitzer et al: The <12 month olds, were divided into monthy age classes. The remaining population was divided into 6 classes: 1–4 years old, 5–9 years, 10–19 years, 20–39 years, 40–59-years, and 60+ years old. In this example, the 1-4 year olds are divided into 12 month age classes as well.

```{r}
#T is N time points; N_ages is number of age groups
parmset <- readRDS('./data_and_parms/parmset.rds')
Pop1 <- readRDS('./data_and_parms/pop1.rds') #initial population
B <-  readRDS('./data_and_parms/Birth_rate.rds')
c2 <- readRDS( './data_and_parms/c2.rds') # contact matrix

@@

@@

```{r}
###########################################
# 1/duration of maternal immunity (DAYS)
DurationMatImmunityDays= 112
###########################################

###########################################
# update LRI and hospitalization probability
###########################################
al=21
delta1=c(rep(.40,3),rep(.39,3), rep(.21,3),rep(.20,3),0.16,rep(.14,3),rep(0.05,al-16))#proportion of first infections that are LRI (by age)
delta2=.5*delta1#proportion of second infectious that are LRI 
delta3=.7*delta2#proportion of third infections that are LRI
hosp1=c(.18*rep(.40,3),0.08*rep(.39,3),0.07*rep(.21,3),0.06*rep(.20,3),0.06*0.16,0.05*rep(.14,3),0.02*rep(0.05,al-16))#proportion of first LRIs that are hospitalized
hosp2=.4*hosp1#proportion of second LRIs that are hospitalized
hosp3=c(rep(0,al-2),0.00001,0.00004)#proportion of third LRIs that are hospitalized

@@

```{r}
source('./data_and_parms/simple_model.R') #Read in the model
```

**NOTE: time step here is in months--need to adjust seasonality accordingly**
```{r}
 start_time = 1 # start date (years)
 tmax = nrow(B)
# end_time = 25 # end date (years)
my_times <- seq(start_time, tmax, by = 1) # gives a sequence from start to end
											 # in increments of 1
```

@@

#########################################
#Seasonal components--
#???Should prob be estimated?
b1=0.1830827 #Seasonal amplitude
phi=3.327749 #Seasonal phase shift (weeks;[-26,26] 0=peak @ Jan 1)
#########################################

@@

###########################################
# 1/duration of maternal immunity (DAYS)
DurationMatImmunityDays= 16*7
###########################################

@@

#############################################
#Birth rate (births/person/YEAR
#Matrix: T rows, N_ages columns; columns 2:N_ages all 0s
PerCapitaBirthsYear=B 
#############################################

@@

baseline.txn.rate <- 9.0
q=1
c2=c2

@@

#Save parameters in a list
```{r saveparms}
parms<-list(PerCapitaBirthsYear=PerCapitaBirthsYear,
            DurationMatImmunityDays=DurationMatImmunityDays,
            WidthAgeClassMonth=WidthAgeClassMonth,
            um=um,
            b1=b1,
            phi=phi,
            rho1=rho1,
            rho2=rho2,
            dur.days1=dur.days1,
            dur.days2=dur.days2,
            dur.days3=dur.days3,
            yinit.matrix=yinit.matrix,
            baseline.txn.rate = baseline.txn.rate,
            q=q,
            c=c2,
            sigma1=sigma1,
            sigma2=sigma2,
            sigma3=sigma3,
            time.step='month'
            )
```

## Run the model 
Note here we are just simulating with set parameters, not fitting to data
```{r}
results <- ode(y=yinit.vector, t=my_times,  
               func=simple_model, 
            parms=parms )

@@

```
```{r}
  St <- results[-c(1:300),-1]
   q=1
   b= parms$baseline.txn.rate/(parms$dur.days1/30.44)
   beta=(b/100)/(sum(yinit.matrix)^(1-q))*c2
   b1= parms$b1
   phi=parms$phi
   rho1=parms$rho1
   rho2=parms$rho2
   sigma1=parms$sigma1
   sigma2=parms$sigma2
   sigma3=parms$sigma3
   t0=108
   I1 <- St[,grep('I1', colnames(St))]
   I2 <- St[,grep('I2', colnames(St))]
   I3 <- St[,grep('I3', colnames(St))]
   I4 <- St[,grep('I4', colnames(St))]
   S1 <- St[,grep('S1', colnames(St))]
   S2 <- St[,grep('S2', colnames(St))]
   S3 <- St[,grep('S3', colnames(St))]
   S0 <- St[,grep('S0', colnames(St))]
   lambda1=matrix(0,nrow=t0,ncol=al)#Force of infection
    for (t in 1:t0)
    {lambda1[t,] <- as.vector((1+b1*cos(2*pi*(t-phi*12)/12))*((I1[t,]+rho1*I2[t,]+rho2*I3[t,]+rho2*I4[t,])%*%beta)/sum(St[t,]))}

@@

## fit the model to simulated data in order to estimate parameter sets
```{r}
parmset<-list(PerCapitaBirthsYear=PerCapitaBirthsYear,
            WidthAgeClassMonth=WidthAgeClassMonth,
            um=um,
            rho1=rho1,
            rho2=rho2,
            dur.days1=dur.days1,
            dur.days2=dur.days2,
            dur.days3=dur.days3,
            yinit.matrix=yinit.matrix,
            q=q,
            c=c2,
            sigma1=sigma1,
            sigma2=sigma2,
            sigma3=sigma3,
            time.step='month'
            )

fitmodel <-  function(parameters,dat) {   # takes the parameter values and dataset as inputs
  protrans <- parameters[1] # estimate parameter related to R0 (baseline transmission rate)
  amp <- parameters[2] # estimate parameter related to seasonal amplitude
  trans <- parameters[3] # estimate parameter related to seasonal peak timing
  DMD <- parameters[4] # estimate parameter related to the duration of maternal immunity
  b1 <- exp(amp) #ensure positive 
  baseline.txn.rate <- exp(protrans) #ensure positive
  phi <-  (2*pi*(exp(trans))) / (1+exp(trans)) # transform to its scale in the model
  durx <- exp(DMD) #ensure positive
  durx <<- durx #somehow we need this for the function and model to recognize these parameters
  baseline.txn.rate <<- baseline.txn.rate
  b1 <<- b1
  phi <<- phi

   # Simulate the model with initial conditions and timesteps defined above, and parameter values from function call
  results <- ode(y=yinit.vector, t=my_times,  
               func=simple_model, 
            parms=c(parmset,baseline.txn.rate=baseline.txn.rate,b1,phi,DurationMatImmunityDays=durx))
  t0 <- 108
  results<- tail(results,t0)
  St <- results[,-1]

   b=baseline.txn.rate/(parmset$dur.days1/30.44)
   beta=(b/100)/(sum(yinit.matrix)^(1-1))*c2

   lambda1=matrix(0,nrow=t0,ncol=al)#Force of infection
    for (t in 1:t0)
    {lambda1[t,] <- as.vector((1+b1*cos(2*pi*(t-phi*12)/12))*((St[t,(2*al+1):(3*al)]+rho1*St[t,(4*al+1):(5*al)]+rho2*St[t,(6*al+1):(7*al)]+rho2*St[t,(8*al+1):(9*al)])%*%beta)/sum(St[t,]))}

  H1=matrix(0,nrow=t0,ncol=al)#Number of hospitalizations by age
  for (i in 1:al){
    H1[,i]=hosp1[i]*St[,(al+i)]*lambda1[,i]+hosp2[i]*sigma1*St[,(3*al+i)]*lambda1[,i]+hosp3[i]*sigma2*St[,(5*al+i)]*lambda1[,i]+hosp3[i]*sigma3*St[,(7*al+i)]*lambda1[,i]
    }
  H <- rowSums(H1)

  agedist <- c(sum(colSums(H1)[1:3]),sum(colSums(H1)[4:6]),
            sum(colSums(H1)[7:9]),
            sum(colSums(H1)[10:12]),
            colSums(H1)[13:21])/sum(H1)

   LLall <- sum(dpois(x = dat, lambda =H, log = TRUE)) # fit to timeseries
   LLmulti <- dmultinom(x= agedist,prob = agedist_Sim,log = T) # fit to age distribution

   #prior
   durprior <- dgamma(x=durx,22,5,log=T)

   #total LL
   LL <- LLall+LLmulti+durprior

   return(LL)
}

fitLL <- optim(par = c(2.198522,  -1.638387 , -1.301047 ,  4.619743 ),           # starting values for beta and gamma - you should get the same result no matter which values you choose here
      fn = fitmodel,        # the distance function to optimise
      dat = Hosp_sim,         # the dataset to fit to ("dat" argument is passed to the function specified in fn)
      control = list(fnscale=-1)) # the log likelihood is negative; here we minimize the negative log likelihood
```

>>>>

---
title: "Antibody_Sim"
author: "ZHE ZHENG"
date: '2022-06-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
time <- 1:(12*9) # time series data point
period = 12 # time unit of observation (12 corresponds to months; 52 corresponds to weeks)
AMP = 1.5 # amplitude of antibody seasonality
phase= 1.6 # shift of peak timing at the beginning
#Naturally accuring maternal antibdoy levels/ harmonic term to represent seasonal waves
mab.p <- 13+AMP*cos(2*pi*(time/period-phase))                                 # Maternal antibody level placebo

#Neonatal half-life (days)
Neo.half.time <- 46.9  # Antibody half life neonates

threshold <-  7      # Protection threshold literature
```

Transfer function - obtained from fitting a regression on published data (see Scheltema et al 2018)
```{r}
transfer.function <- function(transferratio.vector, gestational.age, ga.vector){

  exp.model <- lm(log(transferratio.vector) ~ gestational.age) # regression exponential

  summary.exp.model <- summary(exp.model)

  coef.exp.model <- coef(exp.model) #coefficient

  return(coef.exp.model)
}

coef.exp.model <- transfer.function(transferratio.vector = c(0.025, 0.1, 0.5, 0.70, 1.5), 
                                    gestational.age = c(13, 28, 32, 36, 40),
                                    ga.vector = c(1:42)) 
```

Ab levels mother at time t
```{r}
maternal.ab.level <- mab.p

# change to a function of vaccination
# if vac_status==0, meaning that the individual did not receive vaccine; all-or-nothing settings, can change based on vaccine efficacy/ vaccine response later
## exponential function
# maternal.ab.level <- function(vac_status,t.measure, t.vac, 
#                               vac.effect, m.ab.base){
#   if (vac_status==0){m.ab.level <- m.ab.base}
#   
#   if (vac_status==1){
#   t.since.vac <- t.measure*7 - t.vac*7
#   
#   # before week 1
#   if(t.since.vac < 7) {m.ab.level <- m.ab.base}     
#   
#   # exp between 1 and 2
#   if(t.since.vac >= 7 & t.since.vac < 14) {m.ab.level <- m.ab.base * exp((t.since.vac-7)*(log(vac.effect)/7))}  #new!!! 14 days ipv 21
#   
#   # after week 2
#   if(t.since.vac >= 14) {m.ab.level <- (m.ab.base * vac.effect) * (1/2)^((t.since.vac-14)/(mat.half.time))
#   
#   }}
#   
#   return(m.ab.level)
# }

@@

Ab levels at time of birth
```{r}
#### Cord blood level at time of birth ####
cord.blood.ab.level <- function(gest.age, coef.exp.model, m.ab.base){

  transfer <- exp(coef.exp.model[1] + coef.exp.model[2] * gest.age)

  cb.level <- maternal.ab.level* transfer

  return(cb.level)
}

# cord.blood.ab.level <- function(vac_status,gest.age, t.vac, coef.exp.model, vac.effect, m.ab.base){
#   
#   transfer <- exp(coef.exp.model[1] + coef.exp.model[2] * gest.age)
#   
#   if(vac_status==0){
#   cb.level <- maternal.ab.level(gest.age,vac_status,t.measure, t.vac, 
#                              vac.effect, m.ab.base) * transfer}
#   
#   if(vac_status==1){
#   cb.level <- maternal.ab.level(gest.age,vac_status,t.measure, t.vac, 
#                              vac.effect, m.ab.base) * transfer}
#   
#   return(cb.level)
# }
```

Ab levels after birth
```{r}
#### Cord blood level at time of birth ####
ab.level.afterbirth <- function(t, cb.level){

ab.level <- cb.level * (1/2)^(t/half.time)

  return(ab.level)
}

Length of maternal protection

```{r}
#### Age at threshold####

age.function <- function(cb.level,threshold, half.time){

   exp.age <- ifelse((half.time * (log((threshold/cb.level),base = exp(1))))/-log(2, exp(1))>0,(half.time * (log((threshold/cb.level),base = exp(1))))/-log(2, exp(1)),0)
  # expected age at threshold; greater than 0 or at 0

   return(exp.age)
}
```

```{r}
cb.level = cord.blood.ab.level(gest.age = gest.age, coef.exp.model = coef.exp.model, m.ab.base = mab.p)

Timeprotected <- age.function(cb.level =cb.level,threshold = threshold,half.time = Neo.half.time)

plot(maternal.ab.level,type="l",ylim = c(min(cb.level),max(maternal.ab.level)))
lines(cb.level,col="red")

plot(Timeprotected,type = "l")
```

```{r}
library(deSolve)
library(RColorBrewer)
library(reshape2)
library(dplyr)
library(stringr)
```
In Pitzer et al: The <12 month olds, were divided into monthy age classes. The remaining population was divided into 6 classes: 1–4 years old, 5–9 years, 10–19 years, 20–39 years, 40–59-years, and 60+ years old. In this example, the 1-4 year olds are divided into 12 month age classes as well. 
```{r}
#T is N time points; N_ages is number of age groups
parmset <- readRDS('/Users/zhezheng/Library/CloudStorage/Box-Box/aim3/RSVtransmissionmodel/RSV_metapop/SimpleModel/data_and_parms/parmset.rds')
Pop1 <- readRDS('/Users/zhezheng/Library/CloudStorage/Box-Box/aim3/RSVtransmissionmodel/RSV_metapop/SimpleModel/data_and_parms/pop1.rds') #intial population
NL_B <- read.csv("~/Library/CloudStorage/OneDrive-YaleUniversity/SPREAD study/SPREAD/NL_birthrate.csv")
NL_B <- NL_B[complete.cases(NL_B),]
NL_B$birthrate <- NL_B$birth/NL_B$Pop
c2 <- readRDS( '/Users/zhezheng/Library/CloudStorage/Box-Box/aim3/RSVtransmissionmodel/RSV_metapop/SimpleModel/data_and_parms/c2.rds') # contact matrix
```

@@

```{r}
###########################################
# update LRI and hospitalization probability
###########################################
al=21
delta1=c(rep(.40,3),rep(.39,3), rep(.21,3),rep(.20,3),0.16,rep(.14,3),rep(0.05,al-16))#proportion of first infections that are LRI (by age)
delta2=.5*delta1#proportion of second infectious that are LRI 
delta3=.7*delta2#proportion of third infections that are LRI
hosp1=c(.18*rep(.40,3),0.08*rep(.39,3),0.07*rep(.21,3),0.06*rep(.20,3),0.06*0.16,0.05*rep(.14,3),0.02*rep(0.05,al-16))#proportion of first LRIs that are hospitalized
hosp2=.4*hosp1#proportion of second LRIs that are hospitalized
hosp3=c(rep(0,al-2),0.00001,0.00004)#proportion of third LRIs that are hospitalized

@@

```{r}
source("~/Library/CloudStorage/Box-Box/immuno-epi/Rcode/immuno_epi_model.R") #Read in the model
```

**NOTE: time step here is in months--need to adjust seasonality accordingly**
```{r}
 start_time = 1 # start date (years)
 tmax = nrow((NL_B))
# end_time = 25 # end date (years)
my_times <- seq(start_time, tmax, by = 1) # gives a sequence from start to end
											 # in increments of 1
```

@@

#########################################
#Seasonal components--
#???Should prob be estimated?
b1=0.16 #Seasonal amplitude
phi=3.25 #Seasonal phase shift (weeks;[-26,26] 0=peak @ Jan 1)
#########################################

@@

###########################################
# 1/duration of maternal immunity (DAYS)
DurationMatImmunityDays= c(rep(0,4),rep(Timeprotected,nrow(NL_B)/length(Timeprotected)))
###########################################

@@

#############################################
#Birth rate (births/person/YEAR
#Matrix: T rows, N_ages columns; columns 2:N_ages all 0s
PerCapitaBirthsMonth=matrix(c(NL_B$birthrate,rep(0,nrow(NL_B)*(nrow(yinit.matrix)-1))),byrow = F,nrow=nrow(NL_B),ncol=nrow(yinit.matrix))

PerCapitaBirthsMonth_total=rbind(do.call(rbind, replicate(6, PerCapitaBirthsMonth[7:18,], simplify=FALSE)),PerCapitaBirthsMonth[19:24,])
PerCapitaBirthsMonth_total=rbind(PerCapitaBirthsMonth_total,PerCapitaBirthsMonth)
PerCapitaBirthsMonth_total=rbind(PerCapitaBirthsMonth_total,PerCapitaBirthsMonth[327:328,])
#############################################

@@

baseline.txn.rate <-9
q=1
c2=c2

@@

#Save parameters in a list
```{r saveparms}
parms<-list(PerCapitaBirthsMonth=PerCapitaBirthsMonth_total,
            DurationMatImmunityDays=DurationMatImmunityDays,
            WidthAgeClassMonth=WidthAgeClassMonth,
            um=um,
            b1=b1,
            phi=phi,
            rho1=rho1,
            rho2=rho2,
            dur.days1=dur.days1,
            dur.days2=dur.days2,
            dur.days3=dur.days3,
            yinit.matrix=yinit.matrix,
            baseline.txn.rate = baseline.txn.rate,
            q=q,
            c=c2,
            sigma1=sigma1,
            sigma2=sigma2,
            sigma3=sigma3,
            time.step='month',
            cocooning_duration=7
            )
```

## Run the model 
Note here we are just simulating with set parameters, not fitting to data
```{r}
results <- ode(y=yinit.vector, t=my_times,  
               func=immuno_epi_model, 
            parms=parms )

@@

```
```{r}
  St <- results[-c(1:220),-1]
   q=1
   b= parms$baseline.txn.rate/(parms$dur.days1/30.44)
   beta=(b/100)/(sum(yinit.matrix)^(1-q))*c2
   b1= parms$b1
   phi=parms$phi
   rho1=parms$rho1
   rho2=parms$rho2
   sigma1=parms$sigma1
   sigma2=parms$sigma2
   sigma3=parms$sigma3
   t0=108
   I1 <- St[,grep('I1', colnames(St))]
   I2 <- St[,grep('I2', colnames(St))]
   I3 <- St[,grep('I3', colnames(St))]
   I4 <- St[,grep('I4', colnames(St))]
   S1 <- St[,grep('S1', colnames(St))]
   S2 <- St[,grep('S2', colnames(St))]
   S3 <- St[,grep('S3', colnames(St))]
   S0 <- St[,grep('S0', colnames(St))]
   lambda1=matrix(0,nrow=t0,ncol=al)#Force of infection
    for (t in 1:t0)
    {lambda1[t,] <- as.vector((1+b1*cos(2*pi*(t-phi*12)/12))*((I1[t,]+rho1*I2[t,]+rho2*I3[t,]+rho2*I4[t,])%*%beta)/sum(St[t,]))}

@@

### 2023-01-26 5_roadmaps_future_projects/RSV_maternal_vac_eff/Sim_H_vac.Rmd
add materials

---
title: "Antibody_Sim"
author: "ZHE ZHENG"
date: '2022-06-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
time <- 1:(12*9) # time series data point
period = 12 # time unit of observation (12 corresponds to months; 52 corresponds to weeks)
AMP = 1.5 # amplitude of antibody seasonality
phase= 1.6 # shift of peak timing at the beginning
#Naturally accuring maternal antibdoy levels/ harmonic term to represent seasonal waves
mab.p <- 13+AMP*cos(2*pi*(time/period-phase))                                 # Maternal antibody level placebo

#Neonatal half-life (days)
Neo.half.time <- 46.9  # Antibody half life neonates

threshold <-  7      # Protection threshold literature
```

Transfer function - obtained from fitting a regression on published data (see Scheltema et al 2018)
```{r}
transfer.function <- function(transferratio.vector, gestational.age, ga.vector){

  exp.model <- lm(log(transferratio.vector) ~ gestational.age) # regression exponential

  summary.exp.model <- summary(exp.model)

  coef.exp.model <- coef(exp.model) #coefficient

  return(coef.exp.model)
}

coef.exp.model <- transfer.function(transferratio.vector = c(0.025, 0.1, 0.5, 0.70, 1.5), 
                                    gestational.age = c(13, 28, 32, 36, 40),
                                    ga.vector = c(1:42)) 
```

Ab levels mother at time t
```{r}
maternal.ab.level <- mab.p

# change to a function of vaccination
# if vac_status==0, meaning that the individual did not receive vaccine; all-or-nothing settings, can change based on vaccine efficacy/ vaccine response later
## exponential function
# maternal.ab.level <- function(vac_status,t.measure, t.vac, 
#                               vac.effect, m.ab.base){
#   if (vac_status==0){m.ab.level <- m.ab.base}
#   
#   if (vac_status==1){
#   t.since.vac <- t.measure*7 - t.vac*7
#   
#   # before week 1
#   if(t.since.vac < 7) {m.ab.level <- m.ab.base}     
#   
#   # exp between 1 and 2
#   if(t.since.vac >= 7 & t.since.vac < 14) {m.ab.level <- m.ab.base * exp((t.since.vac-7)*(log(vac.effect)/7))}  #new!!! 14 days ipv 21
#   
#   # after week 2
#   if(t.since.vac >= 14) {m.ab.level <- (m.ab.base * vac.effect) * (1/2)^((t.since.vac-14)/(mat.half.time))
#   
#   }}
#   
#   return(m.ab.level)
# }

Ab levels at time of birth
```{r}
#### Cord blood level at time of birth ####
cord.blood.ab.level <- function(gest.age, coef.exp.model, m.ab.base){

  transfer <- exp(coef.exp.model[1] + coef.exp.model[2] * gest.age)

  cb.level <- maternal.ab.level* transfer

  return(cb.level)
}

# cord.blood.ab.level <- function(vac_status,gest.age, t.vac, coef.exp.model, vac.effect, m.ab.base){
#   
#   transfer <- exp(coef.exp.model[1] + coef.exp.model[2] * gest.age)
#   
#   if(vac_status==0){
#   cb.level <- maternal.ab.level(gest.age,vac_status,t.measure, t.vac, 
#                              vac.effect, m.ab.base) * transfer}
#   
#   if(vac_status==1){
#   cb.level <- maternal.ab.level(gest.age,vac_status,t.measure, t.vac, 
#                              vac.effect, m.ab.base) * transfer}
#   
#   return(cb.level)
# }
```

Ab levels after birth
```{r}
#### Cord blood level at time of birth ####
ab.level.afterbirth <- function(t, cb.level){

ab.level <- cb.level * (1/2)^(t/half.time)

  return(ab.level)
}

Length of maternal protection

```{r}
#### Age at threshold####

age.function <- function(cb.level,threshold, half.time){

   exp.age <- ifelse((half.time * (log((threshold/cb.level),base = exp(1))))/-log(2, exp(1))>0,(half.time * (log((threshold/cb.level),base = exp(1))))/-log(2, exp(1)),0)
  # expected age at threshold; greater than 0 or at 0

   return(exp.age)
}
```

```{r}
cb.level = cord.blood.ab.level(gest.age = gest.age, coef.exp.model = coef.exp.model, m.ab.base = mab.p)

Timeprotected <- age.function(cb.level =cb.level,threshold = threshold,half.time = Neo.half.time)

plot(maternal.ab.level,type="l",ylim = c(min(cb.level),max(maternal.ab.level)))
lines(cb.level,col="red")

plot(Timeprotected,type = "l")
```

@@

#View contact matrix
```{r}
n.cols=100
nice.cols <-  colorRampPalette(brewer.pal(9, "YlOrRd"))(n.cols)
heatmap(c2/sum(diag(c2)), Rowv=NA, Colv=NA, scale='none', col=nice.cols)
```
Set names of Age groups 
```{r}
N_ages <- length(Pop1) 
agenames <- paste0('Agegrp', 1:N_ages) #Could replace this with vector of actual age names
```

@@

#########################################
#Seasonal components--
#???Should prob be estimated?
b1=0.16 #Seasonal amplitude
phi=3.25 #Seasonal phase shift (weeks;[-26,26] 0=peak @ Jan 1)
#########################################

@@

###########################################
# 1/duration of maternal immunity (DAYS)
DurationMatImmunityDays= c(rep(0,4),rep(Timeprotected,nrow(NL_B)/length(Timeprotected)))
###########################################

@@

##To ESTIMTE from DATA?: Baseline transmission rate
#############################################

baseline.txn.rate <-9
q=1
c2=c2

@@

#Save parameters in a list
```{r saveparms}
parms<-list(PerCapitaBirthsMonth=PerCapitaBirthsMonth_total,
            DurationMatImmunityDays=DurationMatImmunityDays,
            WidthAgeClassMonth=WidthAgeClassMonth,
            um=um,
            b1=b1,
            phi=phi,
            rho1=rho1,
            rho2=rho2,
            dur.days1=dur.days1,
            dur.days2=dur.days2,
            dur.days3=dur.days3,
            yinit.matrix=yinit.matrix,
            baseline.txn.rate = baseline.txn.rate,
            q=q,
            c=c2,
            sigma1=sigma1,
            sigma2=sigma2,
            sigma3=sigma3,
            time.step='month',
            cocooning_duration=7
            )
```

## Run the model 
Note here we are just simulating with set parameters, not fitting to data
```{r}
results <- ode(y=yinit.vector, t=my_times,  
               func=immuno_epi_model, 
            parms=parms )

## Extract the relevant results
```{r}

#Pitzer used a 40-50 YEAR burn in period
burnN <- 20*12
results.burned <- results[-c(1:burnN),]

##Any infected person
infected.cols <- results.burned[,c(grep('I1', colnames(results.burned)),grep('I2', colnames(results.burned)),grep('I3', colnames(results.burned)),grep('I4', colnames(results.burned)  )) ]

infected.all <- apply(infected.cols,1, sum)

infected.cols.m <- melt(infected.cols)
infected.cols.m$agegrp <- sub(" .*", "",infected.cols.m$Var2 )
infected.cols.c <- dcast(infected.cols.m, Var1~agegrp, fun.aggregate = sum)

#Pop
all.m <- melt(results.burned[,grep('Agegrp',colnames(results.burned))])
all.m$agegrp <- sub(" .*", "",all.m$Var2 )
all.c <- dcast(all.m, Var1~agegrp, fun.aggregate = sum)
pop.all <- apply(all.c[,-1],1,sum)

plot(infected.all, type='l')
plot(pop.all,type="l")

```
```{r}
  St <- results[-c(1:220),-1]
   q=1
   b= parms$baseline.txn.rate/(parms$dur.days1/30.44)
   beta=(b/100)/(sum(yinit.matrix)^(1-q))*c2
   b1= parms$b1
   phi=parms$phi
   rho1=parms$rho1
   rho2=parms$rho2
   sigma1=parms$sigma1
   sigma2=parms$sigma2
   sigma3=parms$sigma3
   t0=108
   I1 <- St[,grep('I1', colnames(St))]
   I2 <- St[,grep('I2', colnames(St))]
   I3 <- St[,grep('I3', colnames(St))]
   I4 <- St[,grep('I4', colnames(St))]
   S1 <- St[,grep('S1', colnames(St))]
   S2 <- St[,grep('S2', colnames(St))]
   S3 <- St[,grep('S3', colnames(St))]
   S0 <- St[,grep('S0', colnames(St))]
   lambda1=matrix(0,nrow=t0,ncol=al)#Force of infection
    for (t in 1:t0)
    {lambda1[t,] <- as.vector((1+b1*cos(2*pi*(t-phi*12)/12))*((I1[t,]+rho1*I2[t,]+rho2*I3[t,]+rho2*I4[t,])%*%beta)/sum(St[t,]))}

@@

```{r}
plot(1:108,rowSums(H1),type="l",xlab="months",ylab="simulated RSV hospitalizations")
# need to get rid of the burn-in period

barplot(c(sum(colSums(H1)[1:3]),sum(colSums(H1)[4:6]),
            sum(colSums(H1)[7:9]),
            sum(colSums(H1)[10:12]),
            colSums(H1)[13:21])/sum(H1),xlab="age group",ylab="Number of Hospitalizations")

# simulated age distribution Age groups: The <12 month olds, were divided into per 3-months age classes. The 1-4 year olds are  divided into 1yearly age classes, then 5–9 years, 10–19 years, 20–39 years, 40–59-years, and 60+ years old

Hosp_sim <- as.integer(rowSums(H1))
agedist_Sim <- c(sum(colSums(H1)[1:3]),sum(colSums(H1)[4:6]),
            sum(colSums(H1)[7:9]),
            sum(colSums(H1)[10:12]),
            colSums(H1)[13:21])/sum(H1)
```

>>>>

---
title: "simulate Hospitalizations"
author: "ZHE ZHENG"
date: '2022-11-04'
output: html_document
---

@@

#View contact matrix
```{r}
n.cols=100
nice.cols <-  colorRampPalette(brewer.pal(9, "YlOrRd"))(n.cols)
heatmap(c2/sum(diag(c2)), Rowv=NA, Colv=NA, scale='none', col=nice.cols)
```

Set names of Age groups 
```{r}
N_ages <- length(Pop1) 
agenames <- paste0('Agegrp', 1:N_ages) #Could replace this with vector of actual age names
```

@@

#########################################
#Seasonal components--
#???Should prob be estimated?
b1=0.2 #Seasonal amplitude
phi=3.65 #Seasonal phase shift (weeks;[-26,26] 0=peak @ Jan 1)
#########################################

@@

###########################################
# 1/duration of maternal immunity (DAYS)
DurationMatVac= 60
DurationNaturalProtection= 30
###########################################

@@

#############################################
# frequency dependent or density dependent transmission
q=1

#############################################
#contact matrix
c2=c2

##Baseline transmission rate determine the force of transmission
#############################################

baseline.txn.rate <-9

@@

#Save parameters in a list
```{r saveparms}
parms<-list(PerCapitaBirthsYear=PerCapitaBirthsMonth_total*12,
            DurationMatVac=DurationMatVac,
            WidthAgeClassMonth=WidthAgeClassMonth,
            um=um,
            b1=b1,
            phi=phi,
            rho1=rho1,
            rho2=rho2,
            dur.days1=dur.days1,
            dur.days2=dur.days2,
            dur.days3=dur.days3,
            yinit.matrix=yinit.matrix,
            baseline.txn.rate = baseline.txn.rate,
            q=q,
            c=c2,
            sigma1=sigma1,
            sigma2=sigma2,
            sigma3=sigma3,
            time.step='month',
            DurationNaturalProtection=DurationNaturalProtection
            )
```

## Run the model 
Note here we are just simulating with set parameters, not fitting to data
```{r}
results <- ode(y=yinit.vector, t=my_times,  
               func=immuno_epi_model, 
            parms=parms )
```

## Get hospitalizations based on FOI
```{r}
  St <- results[-c(1:220),-1]
   q=1
   b= parms$baseline.txn.rate/(parms$dur.days1/30.44)
   beta=(b/100)/(sum(yinit.matrix)^(1-q))*c2
   b1= parms$b1
   phi=parms$phi
   rho1=parms$rho1
   rho2=parms$rho2
   sigma1=parms$sigma1
   sigma2=parms$sigma2
   sigma3=parms$sigma3
   t0=108
   I1 <- St[,grep('I1', colnames(St))]
   I2 <- St[,grep('I2', colnames(St))]
   I3 <- St[,grep('I3', colnames(St))]
   I4 <- St[,grep('I4', colnames(St))]
   S1 <- St[,grep('S1', colnames(St))]
   S2 <- St[,grep('S2', colnames(St))]
   S3 <- St[,grep('S3', colnames(St))]
   S0 <- St[,grep('S0', colnames(St))]
   lambda1=matrix(0,nrow=t0,ncol=al)#Force of infection
    for (t in 1:t0)
    {lambda1[t,] <- as.vector((1+b1*cos(2*pi*(t-phi*12)/12))*((I1[t,]+rho1*I2[t,]+rho2*I3[t,]+rho2*I4[t,])%*%beta)/sum(St[t,]))}

@@

Hosp_under5 <- H1[(nrow(H1)-11):nrow(H1),1:16]

# number of hospitalizations depends on the season of observation 
# (birth to 90 days)
Hosp_vac_90 <- Hosp_90(t=101,H1)
# number of hospitalizations depends on the season of observation 
# (birth to 180 days)
#Hosp_vac_180 <- Hosp_180(t=90,H1)

# Vaccine efficacy 1-Hosp_mat_vac/Hosp_no_vac
1-Hosp_vac_90/Hosp_novac_90
```

## calculate the age of hospitalizations (age unit: month)
```{r}
age_midpoint <- c(seq(1/24,23/24,1/12),1.5,2.5,3.5,4.5)*12
time.step="month"

average_age_H(age_midpoint,Hosp_under5,time.step)
```
